Homework 4 

Joel McClure

1. A maker of dietary fiber supplement tablets claims its bottles contain, on average, 250 g of the supplement,with a standard deviation of 2 g.

(a) Find the probability a randomly selected bottle weighs between 246 and 254 g. (To do this, suppose
the bottle weights are approximately normally distributed.)

```{r}
pnorm(254,250,2)-pnorm(246,250,2)
```


(b) 100 bottles are selected randomly. Find the probability that their average weight is less than 250 g.(Do not assume any longer that the bottle weights are normally distributed.)


```{r}
cat("N > 30, we will use the central limit theorem",
"\nz = ((250-250)/(2^2/sqrt(100)))","\nz = 0","\nP(AWB) = 0.5","\nP(AWB) =", pnorm(250,250,2))
```


2. The Hereford Cattle Society says that the mean weight of a one-year-old Hereford bull is 1135 pounds, with a standard deviation of 97 pounds. Suppose 40 bulls are randomly selected and loaded on a train car. Find the probability their combined weight exceeds 46000 pounds.

```{r}
x = 1135 * 40 
vr = 40^2 * 97^2
sd = sqrt(vr)
z = ((46000-x)/(sd/sqrt(40)))
z = 0.978
cat("P(CWEP) = 0.1635", "\nP(CWEP) =", pnorm(1150, 1135,sqrt(97^2/40),FALSE))
```

3. Suppose each of the following data sets is a simple random sample from some population. For each data set, make a normal QQ plot. Look at the QQ plot and say whether the data are plausibly from a normal population or whether they are evidence against a normal population. 

```{r}
#Here are reference plots
#for (i in 1:12){
 # qqnorm(rnorm(length(d3a)))}
```


(a) data = c(7,13.2,8.1,8.2,6,9.5,9.4,8.7,9.8,10.9,8.4,7.4,8.4,10,9.7,8.6,12.4,10.7,11,9.4)
```{r}
d3a = c(7,13.2,8.1,8.2,6,9.5,9.4,8.7,9.8,10.9,8.4,7.4,8.4,10,9.7,8.6,12.4,10.7,11,9.4)
qqnorm(d3a)
```
This looks to plausibly be from a normal distribution. The datapoints appear to be very close to where a linear trendline would be. 

(b) data = c(2.5,1.8,2.6,-1.9,1.6,2.6,1.4,0.9,1.2,2.3,-1.5,1.5,2.5,2.9,-0.1)
```{r}
d3b = c(2.5,1.8,2.6,-1.9,1.6,2.6,1.4,0.9,1.2,2.3,-1.5,1.5,2.5,2.9,-0.1)
qqnorm(d3b)
```
This plot varies from the trendline by quite a bit over the points. However this amount of variation is shown in our sample normal distributions so it is plausable that it could be from a normal distribution even though I lack confidence in that answer. 

(c) data = c(3.3,1.7,3.3,3.3,2.4,0.5,1.1,1.7,12,14.4,12.8,11.2,10.9,11.7,11.7,11.6)
```{r}
d3c = c(3.3,1.7,3.3,3.3,2.4,0.5,1.1,1.7,12,14.4,12.8,11.2,10.9,11.7,11.7,11.6)
qqnorm(d3c)
```
This plot appears to have two groups of data that could show a binomial distribution. This data does not appear to have a normal distribution due to the break in the middle and the potentail for a binomial distribution 

4. Biologists studying the healing of skin wounds measured the rate at which new cells closed a cut made in the skin of an anesthetized newt. Here are data from a random sample of 18 newts, measured in micrometers (millionths of a meter) per hour: 29, 27, 34, 40, 22, 28, 14, 35, 26, 35, 12, 30, 23, 18, 11, 22, 23, 33

(a) Create a QQ plot of the data. Do you think it is reasonable to assume that the population distribution is normal? Explain your answer. (There isn’t a unique “right” answer.)
```{r}
d4a = c(29, 27, 34, 40, 22, 28, 14, 35, 26, 35, 12, 30, 23, 18, 11, 22, 23, 33)
qqnorm(d4a)
```
This qqplot indicates that a normal distribution is plausible. Aside from the first two points, the data tend to follow a linear line. I would like to see more data to show if the first two points are due to noise or are indicitive of a non-normal trend in the data. 

(b) Regardless of your answer to (a), assume the population distribution is normal and use that assumption to create a 90% CI for µ. (Use a calculator to find ¯x and s and then use the formula provided in class to prepare for exams. Then check your answer with R if you wish.)

```{r}
n = length(d4a)
a = .1
t = 1.74
x = mean(d4a)
s = sd(d4a)
s
low = x-(t*(s/sqrt(n)))
high =  x+(t*(s/sqrt(n)))
cat("We are 90% sure that the true mean value is between",low,"and", high)
```


(c) What sample size n is required to get a 90% confidence interval with error margin 2.0?

```{r}
z=1.645
m=2.0
n = (z*s/m)^2
cat("The sample size",round(n,0),"is sufficent for us to get an error magin of 2.0")
```



5. A pumpkin farmer weighed a simple random sample of size n = 20 pumpkins, with these results: 9.6, 8.8, 5.1, 9.7, 9.1, 8.9, 8, 9.2, 2.7, 9.1, 8.5, 7.3, 9.3, 9.6, 4.1, 9.9, 7.6, 9, 7.2, 8.5. Create a QQ plot of the weights. Do you think it is reasonable to assume that the population distribution is normal? Explain your answer. (There isn’t a unique “right” answer to this problem.)


```{r}
d5 = c(9.6, 8.8, 5.1, 9.7, 9.1, 8.9, 8, 9.2, 2.7, 9.1, 8.5, 7.3, 9.3, 9.6, 4.1, 9.9, 7.6, 9, 7.2, 8.5)
qqnorm(d5)
```
This sample does not show that the population has a normal distribution. The majority of the points appear to be above where the linear trendline would be which would show a left skew. This is just a small sample of the overlying population and a larger sample could very well fix the skew that is showing up in this sample. 